Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-17 Paper.tex
More timestamp fixes.
Fixes for Marko's bugs.

Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
& \wedge & H_d = D(H) \\
& \wedge & H_t > {P(H)_H}_t
\end{eqnarray}

@@

Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
\begin{equation}
(\mathbb{S}_P, g') \equiv \begin{cases}
\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
\end{cases}
\end{equation}

@@

Finally, the contract is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}). Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made. As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}'$ and available gas remaining $g'$.

@@

If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere. Thus formally, we may specify the new state and gas as $\mathbb{S}'$, $g'$ where:
\begin{eqnarray}
(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
\mathbb{S}_P & \equiv & \begin{cases}
\mathbb{S}_0 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
I_a & \equiv & a \\
I_o & \equiv & s \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & [] \\
I_s & \equiv & s \\
I_v & \equiv & v \\
I_\mathbf{b} & \equiv & \mathbf{i}
\end{eqnarray}

@@

If the recipient account contains no code to be executed (i.e. it's not a contract), then we define the final state as being equivalent to the rollback state. In the case that the recipient contains code to be executed (i.e. that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_{code}]$) is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
\begin{eqnarray}
\mathbb{S}' & \equiv & \begin{cases}
\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
I_a & \equiv & a \\
I_o & \equiv & o \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & \mathbf{d} \\
I_s & \equiv & s \\
I_v & \equiv & v \\
\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
\end{eqnarray}

@@

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
\end{eqnarray}

@@

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_S[1] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 1) \div 32 })$ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\
0 & \text{if} \quad \mathbb{M}_S[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_S[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}

>>>>

Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
& \wedge & H_d = D(H) \\
& \wedge & H_s > {P(H)_H}_s
\end{eqnarray}

@@

Evaluating $\mathbb{S}_P$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
\begin{equation}
(\mathbb{S}_P, g') \equiv \begin{cases}
\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
\end{cases}
\end{equation}

@@

Finally, the contract is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}). Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made. As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}^{**}$ and available gas remaining $g'$.

@@

If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere. Thus formally, we may specify the resultant state and gas as $\mathbb{S}_P$, $g'$ where:
\begin{eqnarray}
(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
\mathbb{S}_P & \equiv & \begin{cases}
\mathbb{S}_0 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
I_a & \equiv & a \\
I_o & \equiv & s \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & [] \\
I_s & \equiv & s \\
I_v & \equiv & v \\
I_\mathbf{b} & \equiv & \mathbf{i}
\end{eqnarray}

@@

If the recipient account contains no code to be executed (i.e. it's not a contract), then we define the final state as being equivalent to the rollback state. In the case that the recipient contains code to be executed (i.e. that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_{code}]$) is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
\begin{eqnarray}
\mathbb{S}_P & \equiv & \begin{cases}
\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
\mathbb{S}^{**} & \text{otherwise}
\end{cases} \\
(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
I_a & \equiv & a \\
I_o & \equiv & o \\
I_p & \equiv & p \\
I_\mathbf{d} & \equiv & \mathbf{d} \\
I_s & \equiv & s \\
I_v & \equiv & v \\
\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
\end{eqnarray}

@@

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, S)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

@@

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_T\rVert + \Delta \\
\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_S[x] & \equiv & \mathbb{M}_S[x+\Delta]
\end{eqnarray}

@@

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below. Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.

\subsection{Instruction Set}

Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$. For each instruction, also specify is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack.

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_S[1] \equiv \mathbb{M}_S[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 1) \div 32 })$ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
&&&& $C_{\text{\tiny SSTORE}}(\mathbb{S}, \mathbb{M}) \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\
0 & \text{if} \quad \mathbb{M}_S[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_S[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}
